---
title: "Descriptive BRFSS Healthcare Data Analytics using R"
author: "Nazima Khan"
date: "May 9, 2017"
output: html_document
---
## Objective
#### To Analyze Behavior Risk Factor Surveillance System (BRFSS) for "VETERANS" using BRFSS dataset.

## Brief Summary on the topic
#### What is risk factor surveillance?
#### Keeping track of the rates of risk factors which are the things or states in our daily lives that confers risk to our health is defined as a Risk Factor Surveillance.
#### There are two main Surveillance systems in the United States
* National Health and Nutrition Examination Survey(NHANES).
* Behavior Risk Factor Surveillance System (BRFSS).

### We will be using BRFSS dataset
* BRFSS is a federal and state collaboration
* Data collectors call randomly generated landline phone numbers or cell phone numbers.
* Data from BRFSS are publicly available and can be used widely by CDC for healthcare program planning. And also independent researchers can use this data for in-depth reseacrch and analytics purpose.

###Types of BRFSS Analytics
#### Descriptive Analysis
* Aimed at developing population-based rates.
* Dependent upon the sampling approach--uses "weights".
* More often done by CDC and states.

#### Cross-Sectional Analysis
* Aimed at exploring cross-sectional associations (hinting to potential causes)
* Weighting is generally not used.
* More often done by independent researchers.

### Resources
#### BRFSS Resource Provided by CDC.
* Datasets https://www.cdc.gov/brfss/annual_data/annual_data.htm
* Questionnaires.
* Codebook - Provides univariate summary statistics about each variable.
* Documentation and explanations.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Environment Setup
```{r,include=TRUE,warning=FALSE,message=FALSE}

# Load the required packages (if packages are not available, install them first)
for (package in c('foreign','gtools','questionr','MASS','caret','readr','ggplot2','magrittr','ggthemes','dplyr','corrplot','caTools')) {
  if (!require(package, character.only=T, quietly=T)) {
    install.packages(package)
    library(package,character.only=T)
  }
}
# Package Desription
# foreign   : Reads in "foreign" data types.
# gtools    : Allows you to make macros.
# dplyr     : Helps packages for calculating means and standard deviations.
# questionr : Allows you to do a weighted analysis.
# MASS      : Allows you to do bivariate tests.
```

### Designing Metadata
```{r}
# Set up the data dictionary 
# Let us assume we are referring our native variables for " US Army Dataset". Therefore, we will refer the data variable in ARMY terms i.e.
# 1. Let's select a subpopulation (veterans)
# 2. Let's select a risk factor (alcohol).
# 3. Let's select two outcomes.
#     Continuous: hours of sleep per night
#     Categorical: asthma(yes/no)
# Define Confounding Variables (variable associated with the Exposure and with the Outcome)
```

### Reading and Cleaning the Data
```{r ,include=TRUE,warning=FALSE,message=FALSE}
##Get the Data 
library(foreign)
# Read the csv file and save in an object called "BRFSS_data"
BRFSS_a <- read.xport("C:/CompleteMLProjects/Healthcare/BRFSS/Analytics/Data/LLCP2014.xpt")
colnames(BRFSS_a)
```

### Subset unwanted values from the dataset
```{r}
#define object list of variables to be kept
BRFSSVarList <- c("VETERAN3", 
			"ALCDAY5",
			"SLEPTIM1",
			"ASTHMA3",
			"X_AGE_G",
			"SMOKE100",
			"SMOKDAY2",
			"SEX",
			"X_HISPANC",
			"X_MRACE1",
			"MARITAL",
			"GENHLTH",
			"HLTHPLN1",
			"EDUCA",
			"INCOME2",
			"X_BMI5CAT",
			"EXERANY2")

# subset by varlist
BRFSS_b <- BRFSS_a[BRFSSVarList]

# check columns
colnames(BRFSS_b)

# check rows
nrow(BRFSS_b)

# SUBSET AND SEPERATE NUMBER OF VETERANS AND NON VETERANS FROM THE DATASET
BRFSS_c <- subset(BRFSS_b,VETERAN3==1)

# Check the variable value 
# BRFSS_c$VETERAN3
# Check the number of rows for BRFSS_c
nrow(BRFSS_c)

# We can see 62120 number of VETERANS are in the dataset
# Also we now know that  464664 - 62120 = 402544 number of non veterans are in the dataset.

# ONLY KEEP ROWS WITH VALID ALCOHOL/EXPOSURE VARIABLE.
BRFSS_d <- subset(BRFSS_c, ALCDAY5 < 777 | ALCDAY5 == 888)

# Take a look at the data
# BRFSS_d$ALCDAY5
# Check the number of rows for BRFSS_d
nrow(BRFSS_d)
# 58991 are the number of veterans that do not consume Alcohol.
# Hence we know now 62120 - 58991 = 3129 number of veterans that consume Alcohol.
 

# EXCLUDING SLEEP TIME VARIABLES
# Only keep variable with valid sleep data
BRFSS_e <- subset(BRFSS_d,SLEPTIM1 < 77)

# Check the number of rows for BRFSS_e
nrow(BRFSS_e)
# 58321 are the number of veterans with valid sleep data.
# Hence we know now 58991 - 58321 = 670 number of veterans that have valid sleep pattern.

# EXCLUDING ASTHMA VARIABLES 

# Only keep variables with valid Asthma data
BRFSS_f <- subset(BRFSS_e, ASTHMA3 < 7)

# Check the number of rows for BRFSS_f
nrow(BRFSS_f)
# 58131 are the number of veterans with valid Asthma data.
# Hence we know now 58321 - 58131 = 190 number of veterans  have valid Asthma data.
```

### Generating Exposure Variable
#### First, we will go to our exposure, alcohol. Make a grouping variable for alcohol, and indicator variables for drinking monthly and drinking weekly
#### From the Data Dictionary on the ALCDAY5 tab, we see that if ALCDAY5 falls in this range, 101 to 199, our ALCGRP variable should be coded as three, drink weekly. And those in the 201 to 299 range get a two for drink monthly. And the 888's get a one for no drinks. And the rest are nine, for unknown.
#### Also If ALCGRP is two, the drink monthly flag will be one and everyone else gets a zero. If ALCGRP is three, the drink weekly flag will be one and everyone else gets a zero.
```{r}
# Add Indicator variable for Veterans
# First make copy of the dataset
BRFSS_g <- BRFSS_f

# add the categorical variable set to 9 to the dataset
BRFSS_g$ALCGRP <- 9

# update according to data Dictionary
BRFSS_g$ALCGRP[BRFSS_g$ALCDAY5 < 200] <- 3
BRFSS_g$ALCGRP[BRFSS_g$ALCDAY5 >= 200 & BRFSS_g$ALCDAY5 <777] <- 2
BRFSS_g$ALCGRP[BRFSS_g$ALCDAY5 == 888] <- 1

# Check the variable 
table(BRFSS_g$ALCGRP, BRFSS_g$ALCDAY5)

# Add flags
# Flags for Monthly drinkers
BRFSS_g$DRKMONTHLY <- 0
BRFSS_g$DRKMONTHLY[BRFSS_g$ALCGRP == 2] <- 1

table(BRFSS_g$ALCGRP,BRFSS_g$DRKMONTHLY)

# Flags for Weekly drinkers
BRFSS_g$DRKWEEKLY <- 0
BRFSS_g$DRKWEEKLY[BRFSS_g$ALCGRP == 1] <- 1

table(BRFSS_g$ALCGRP,BRFSS_g$DRKWEEKLY)
```

### Generate outcome variables from data dictionary
#### First, we are going to clean up our outcome variable for sleep duration. Next, we will make sure we have binary variable or flag that is valid for our asthma outcome.
```{r}
#  We need to remove the rows with no information on sleep time and we want to turn our asthma variable into an indicator variable with only ones and zeroes.

# First make copy of the dataset
BRFSS_h <- BRFSS_g

# Make and test sleep variable
# First generate a SLEEPTIM2 variable that is a continuous variable for SLEEEPTIM1 and assign them all as NA
BRFSS_h$SLEPTIM2 <- NA

# Add and check for criteria that SLEPTIM1 cannot be NA and it cannot be 77 and cannot be 99.
BRFSS_h$SLEPTIM2[!is.na(BRFSS_h$SLEPTIM1) & BRFSS_h$SLEPTIM1 !=77 & BRFSS_h$SLEPTIM1 !=99] <- BRFSS_h$SLEPTIM1

# Check the variable
table(BRFSS_h$SLEPTIM1,BRFSS_h$SLEPTIM2)
```

### Make and test asthma variable
```{r}
# Assign 9 to ASTHMA4 
BRFSS_h$ASTHMA4 <- 9

# Then assign 1 to all who have reported ASTHMA
BRFSS_h$ASTHMA4[BRFSS_h$ASTHMA3 == 1] <- 1

# Then assign 0 to all who have reported ASTHMA
BRFSS_h$ASTHMA4[BRFSS_h$ASTHMA3 == 2] <- 0

# Check the variable
table(BRFSS_h$ASTHMA3,BRFSS_h$ASTHMA4)

```

### Generating Age variables
```{r}
# First make copy of the dataset
BRFSS_i <- BRFSS_h

# From the data dictionary by default set the value of all age groups to 0.
# Age group 18 to 24 we can keep it as reference group
BRFSS_i$AGE2 <- 0   # Age 25 to 34
BRFSS_i$AGE3 <- 0   # Age 35 to 44
BRFSS_i$AGE4 <- 0   # Age 45 to 54
BRFSS_i$AGE5 <- 0   # Age 55 to 64
BRFSS_i$AGE6 <- 0   # Age 65 and older

# set conditions to update the flags

BRFSS_i$AGE2[BRFSS_i$X_AGE_G == 2] <- 1
table(BRFSS_i$X_AGE_G,BRFSS_i$AGE2)

BRFSS_i$AGE3[BRFSS_i$X_AGE_G == 3] <- 1
table(BRFSS_i$X_AGE_G,BRFSS_i$AGE3)

BRFSS_i$AGE4[BRFSS_i$X_AGE_G == 4] <- 1
table(BRFSS_i$X_AGE_G,BRFSS_i$AGE4)

BRFSS_i$AGE5[BRFSS_i$X_AGE_G == 5] <- 1
table(BRFSS_i$X_AGE_G,BRFSS_i$AGE5)

BRFSS_i$AGE6[BRFSS_i$X_AGE_G == 6] <- 1
table(BRFSS_i$X_AGE_G,BRFSS_i$AGE6)

```

### MAKE SMOKING VARIABLES
```{r}
# Make smoking variables
BRFSS_i$NEVERSMK <- 0
BRFSS_i$NEVERSMK [BRFSS_i$SMOKE100 == 2] <- 1
table(BRFSS_i$SMOKE100,BRFSS_i$NEVERSMK)

# Make grouping variable
BRFSS_i$SMOKGRP <- 9
BRFSS_i$SMOKGRP[BRFSS_i$SMOKDAY2 == 1 | BRFSS_i$SMOKDAY2 == 2] <- 1
BRFSS_i$SMOKGRP[BRFSS_i$SMOKDAY2 == 3 | BRFSS_i$NEVERSMK == 1] <- 2

table(BRFSS_i$SMOKGRP,BRFSS_i$SMOKDAY2)
table(BRFSS_i$SMOKGRP,BRFSS_i$SMOKE100)

BRFSS_i$SMOKER <- 0
BRFSS_i$SMOKER[BRFSS_i$SMOKGRP == 1] <- 1

table(BRFSS_i$SMOKGRP, BRFSS_i$SMOKER)
```

### Make Sex variable
```{r}
BRFSS_i$MALE <- 0
BRFSS_i$MALE[BRFSS_i$SEX == 1] <- 1

table(BRFSS_i$MALE, BRFSS_i$SEX)
```

### Make Hispanic variable
```{r}
BRFSS_i$HISPANIC <- 0
BRFSS_i$HISPANIC[BRFSS_i$X_HISPANC == 1] <- 1

table(BRFSS_i$HISPANIC, BRFSS_i$X_HISPANC)
```

### Make Race variables
```{r}
BRFSS_i$RACEGRP <- 9
BRFSS_i$RACEGRP[BRFSS_i$X_MRACE1 == 1] <- 1
BRFSS_i$RACEGRP[BRFSS_i$X_MRACE1 == 2] <- 2
BRFSS_i$RACEGRP[BRFSS_i$X_MRACE1 == 3] <- 3
BRFSS_i$RACEGRP[BRFSS_i$X_MRACE1 == 4] <- 4
BRFSS_i$RACEGRP[BRFSS_i$X_MRACE1 == 5] <- 5
BRFSS_i$RACEGRP[BRFSS_i$X_MRACE1 == 6 | BRFSS_i$X_MRACE1 == 7] <- 6

table(BRFSS_i$RACEGRP , BRFSS_i$X_MRACE1)

BRFSS_i$BLACK <- 0
BRFSS_i$ASIAN <- 0
BRFSS_i$OTHRACE <- 0

BRFSS_i$BLACK[BRFSS_i$RACEGRP == 2] <- 1
table(BRFSS_i$RACEGRP, BRFSS_i$BLACK)

BRFSS_i$ASIAN[BRFSS_i$RACEGRP == 4] <- 1
table(BRFSS_i$RACEGRP, BRFSS_i$ASIAN)

BRFSS_i$OTHRACE[BRFSS_i$RACEGRP == 3 | BRFSS_i$RACEGRP == 5 | BRFSS_i$RACEGRP == 6 | BRFSS_i$RACEGRP == 7] <- 1
table(BRFSS_i$RACEGRP, BRFSS_i$OTHRACE)
```

### Make Marital variables
```{r}
BRFSS_i$MARGRP <- 9
BRFSS_i$MARGRP[BRFSS_i$MARITAL == 1 | BRFSS_i$MARITAL == 5] <- 1
BRFSS_i$MARGRP[BRFSS_i$MARITAL == 2 | BRFSS_i$MARITAL == 3 ] <- 2
BRFSS_i$MARGRP[BRFSS_i$MARITAL == 4] <- 3

table(BRFSS_i$MARGRP, BRFSS_i$MARITAL)

BRFSS_i$NEVERMAR <- 0
BRFSS_i$FORMERMAR <- 0

BRFSS_i$NEVERMAR[BRFSS_i$MARGRP == 3] <- 1
table(BRFSS_i$MARGRP, BRFSS_i$NEVERMAR)

BRFSS_i$FORMERMAR[BRFSS_i$MARGRP == 2] <- 1
table(BRFSS_i$MARGRP, BRFSS_i$FORMERMAR)
```

### Make Genhealth variables
```{r}
BRFSS_i$GENHLTH2 <- 9
BRFSS_i$GENHLTH2[BRFSS_i$GENHLTH == 1] <- 1
BRFSS_i$GENHLTH2[BRFSS_i$GENHLTH == 2] <- 2
BRFSS_i$GENHLTH2[BRFSS_i$GENHLTH == 3] <- 3
BRFSS_i$GENHLTH2[BRFSS_i$GENHLTH == 4] <- 4
BRFSS_i$GENHLTH2[BRFSS_i$GENHLTH == 5] <- 5

table(BRFSS_i$GENHLTH2, BRFSS_i$GENHLTH)

BRFSS_i$FAIRHLTH <- 0
BRFSS_i$POORHLTH <- 0

BRFSS_i$FAIRHLTH [BRFSS_i$GENHLTH2 == 4] <- 1
table(BRFSS_i$FAIRHLTH, BRFSS_i$GENHLTH2)

BRFSS_i$POORHLTH [BRFSS_i$GENHLTH2 == 5] <- 1
table(BRFSS_i$POORHLTH, BRFSS_i$GENHLTH2)
```

### Make health plan variables
```{r}
BRFSS_i$HLTHPLN2 <- 9
BRFSS_i$HLTHPLN2[BRFSS_i$HLTHPLN1 == 1] <- 1
BRFSS_i$HLTHPLN2[BRFSS_i$HLTHPLN1 == 2] <- 2

table(BRFSS_i$HLTHPLN1, BRFSS_i$HLTHPLN2)

BRFSS_i$NOPLAN <- 0
BRFSS_i$NOPLAN [BRFSS_i$HLTHPLN2== 2] <- 1
table(BRFSS_i$NOPLAN, BRFSS_i$HLTHPLN2)
```

### Make education variables
```{r}
BRFSS_i$EDGROUP <- 9
BRFSS_i$EDGROUP[BRFSS_i$EDUCA == 1 | BRFSS_i$EDUCA == 2 | BRFSS_i$EDUCA == 3] <- 1
BRFSS_i$EDGROUP[BRFSS_i$EDUCA == 4] <- 2
BRFSS_i$EDGROUP[BRFSS_i$EDUCA == 5] <- 3
BRFSS_i$EDGROUP[BRFSS_i$EDUCA == 6] <- 4

table(BRFSS_i$EDGROUP, BRFSS_i$EDUCA)

BRFSS_i$LOWED <- 0
BRFSS_i$SOMECOLL <- 0

BRFSS_i$LOWED[BRFSS_i$EDGROUP == 1 | BRFSS_i$EDGROUP == 2 ] <- 1
table(BRFSS_i$LOWED, BRFSS_i$EDGROUP)

BRFSS_i$SOMECOLL [BRFSS_i$EDGROUP == 3] <- 1
table(BRFSS_i$SOMECOLL, BRFSS_i$EDGROUP)
```
### Make income variables
```{r}
BRFSS_i$INCOME3 <- BRFSS_i$INCOME2
BRFSS_i$INCOME3[BRFSS_i$INCOME2 >=77] <- 9

table(BRFSS_i$INCOME2, BRFSS_i$INCOME3)

BRFSS_i$INC1 <- 0
BRFSS_i$INC2 <- 0
BRFSS_i$INC3 <- 0
BRFSS_i$INC4 <- 0
BRFSS_i$INC5 <- 0
BRFSS_i$INC6 <- 0
BRFSS_i$INC7 <- 0

BRFSS_i$INC1[BRFSS_i$INCOME3 == 1] <- 1
table(BRFSS_i$INC1, BRFSS_i$INCOME3)

BRFSS_i$INC2[BRFSS_i$INCOME3 == 2] <- 1
table(BRFSS_i$INC2, BRFSS_i$INCOME3)

BRFSS_i$INC3[BRFSS_i$INCOME3 == 3] <- 1
table(BRFSS_i$INC3, BRFSS_i$INCOME3)

BRFSS_i$INC4[BRFSS_i$INCOME3 == 4] <- 1
table(BRFSS_i$INC4, BRFSS_i$INCOME3)

BRFSS_i$INC5[BRFSS_i$INCOME3 == 5] <- 1
table(BRFSS_i$INC5, BRFSS_i$INCOME3)

BRFSS_i$INC6[BRFSS_i$INCOME3 == 6] <- 1
table(BRFSS_i$INC6, BRFSS_i$INCOME3)

BRFSS_i$INC7[BRFSS_i$INCOME3 == 7] <- 1
table(BRFSS_i$INC7, BRFSS_i$INCOME3)
```

### Make BMI variables
```{r}
BRFSS_i$BMICAT<- 9
BRFSS_i$BMICAT[BRFSS_i$X_BMI5CAT ==1] <- 1
BRFSS_i$BMICAT[BRFSS_i$X_BMI5CAT ==2] <- 2
BRFSS_i$BMICAT[BRFSS_i$X_BMI5CAT ==3] <- 3
BRFSS_i$BMICAT[BRFSS_i$X_BMI5CAT ==4] <- 4

table(BRFSS_i$BMICAT, BRFSS_i$X_BMI5CAT)

BRFSS_i$UNDWT <- 0
BRFSS_i$OVWT <- 0
BRFSS_i$OBESE <- 0

BRFSS_i$UNDWT[BRFSS_i$BMICAT== 1] <- 1
table(BRFSS_i$UNDWT, BRFSS_i$BMICAT)

BRFSS_i$OVWT[BRFSS_i$BMICAT== 3] <- 1
table(BRFSS_i$OVWT, BRFSS_i$BMICAT)

BRFSS_i$OBESE[BRFSS_i$BMICAT== 4] <- 1
table(BRFSS_i$OBESE, BRFSS_i$BMICAT)
```

### Make exercise variables
```{r}
BRFSS_i$EXERANY3<- 9
BRFSS_i$EXERANY3[BRFSS_i$EXERANY2 ==1] <- 1
BRFSS_i$EXERANY3[BRFSS_i$EXERANY2 ==2] <- 2

table(BRFSS_i$EXERANY3, BRFSS_i$EXERANY2)

BRFSS_i$NOEXER <- 0
BRFSS_i$NOEXER[BRFSS_i$EXERANY3 ==2] <- 1
table(BRFSS_i$NOEXER, BRFSS_i$EXERANY3)

nrow(BRFSS_i)
```

### Write out analytic dataset
```{r}
write.csv(BRFSS_i, file = "analytic.csv")
```

### Finally now that we have a clean dataset in hand.Let us analyze this data.

```{r}
#read in analytic table
analytic <- read.csv(file="C:/CompleteMLProjects/Healthcare/BRFSS/Analytics/Code/analytic.csv", header=TRUE, sep=",")

#Look at distribution of categorical outcome asthma

AsthmaFreq <- table(analytic$ASTHMA4)
AsthmaFreq
write.csv(AsthmaFreq, file = "AsthmaFreq.csv")

#what proportion of our dataset has ashtma?
PropAsthma <- 5343/52788
PropAsthma

#Look at categorical outcome asthma by exposure, ALCGRP
AsthmaAlcFreq <- table(analytic$ASTHMA4, analytic$ALCGRP)
AsthmaAlcFreq
write.csv(AsthmaAlcFreq, file = "AsthmaAlcFreq.csv")
```

### Look at distribution of sleep duration 
```{r}
#summary statistics
summary(analytic$SLEPTIM2)

#look at histogram and box plot of total file
hist(analytic$SLEPTIM2, 
	main = "Histogram of SLEPTIM2",
	xlab = "Class SLEPTIM2",
	ylab = "Frequency",
	xlim=c(0,15), 
	ylim=c(0,20000),
	border = "red",
	col= "yellow",
	las = 1,
	breaks = 24)

boxplot(analytic$SLEPTIM2, main="Box Plot of SLEPTIM2", 
  	xlab="Total File", ylab="SLEPTIM2")

#See box plots of groups next to each other
boxplot(SLEPTIM2~ALCGRP, data=analytic, main="Box Plot of SLEPTIM2 by ALCGRP", 
  	xlab="ALCGRP", ylab="SLEPTIM2")
```

### Making frequencies per category
```{r}
AsthmaFreq <- table(analytic$ASTHMA4)
AsthmaFreq
write.csv(AsthmaFreq, file = "AsthmaFreq.csv")

AlcFreq <- table(analytic$ALCGRP)
AlcFreq 
write.csv(AlcFreq , file = "AlcFreq.csv")

#USING MACROS

#install package gtools
#then call up library

library(gtools)
#use defmacro to define the macro
FreqTbl <-defmacro(OutputTable, InputVar, CSVTable, 
expr={
OutputTable <- table(InputVar);
write.csv(OutputTable, file = paste0(CSVTable, ".csv"))
})

FreqTbl (AlcFreq, analytic$ALCGRP, "Alc")
FreqTbl (AgeFreq, analytic$X_AGE_G, "Age")
FreqTbl (SexFreq, analytic$SEX, "Sex")
FreqTbl (HispFreq, analytic$X_HISPANC, "Hisp")
FreqTbl (RaceFreq, analytic$RACEGRP, "Race")
FreqTbl (MaritalFreq, analytic$MARGRP, "Mar")
FreqTbl (EdFreq, analytic$EDGROUP, "Ed")
FreqTbl (IncFreq, analytic$INCOME3, "Inc")
FreqTbl (BMIFreq, analytic$BMICAT, "BMI")
FreqTbl (SmokeFreq, analytic$SMOKGRP, "Smok")
FreqTbl (ExerFreq, analytic$EXERANY3, "Exer")
FreqTbl (HlthPlanFreq, analytic$HLTHPLN2, "HlthPln")
FreqTbl (GenHlthFreq, analytic$GENHLTH2, "GenHlth")
```

### Checking for No Asthma frequencies
```{r}
### Subset dataset with only asthma people
asthmaonly <- subset(analytic, ASTHMA4 == 1)
table(asthmaonly$ASTHMA4)
nrow(asthmaonly)

AsthmaFreq <- table(asthmaonly$ASTHMA4)
AsthmaFreq
write.csv(AsthmaFreq, file = "Asthma.csv")

#USING MACROS
library(gtools)
#use defmacro to define the macro
FreqTbl <-defmacro(OutputTable, InputVar, CSVTable, 
expr={
OutputTable <- table(InputVar);
write.csv(OutputTable, file = paste0(CSVTable, ".csv"))
})

FreqTbl (AlcGrpFreq, asthmaonly$ALCGRP, "Alc")
FreqTbl (AgeGrpFreq, asthmaonly$X_AGE_G, "Age")
FreqTbl (SexFreq, asthmaonly$SEX, "Sex")
FreqTbl (HispFreq, asthmaonly$X_HISPANC, "Hisp")
FreqTbl (RaceFreq, asthmaonly$RACEGRP, "Race")
FreqTbl (MaritalFreq, asthmaonly$MARGRP, "Mar")
FreqTbl (EdFreq, asthmaonly$EDGROUP, "Ed")
FreqTbl (IncFreq, asthmaonly$INCOME3, "Inc")
FreqTbl (BMIFreq, asthmaonly$BMICAT, "BMI")
FreqTbl (SmokeFreq, asthmaonly$SMOKGRP, "Smok")
FreqTbl (ExerFreq, asthmaonly$EXERANY3, "Exer")
FreqTbl (HlthPlanFreq, asthmaonly$HLTHPLN2, "HlthPln")
FreqTbl (GenHlthFreq, asthmaonly$GENHLTH2, "GenHlth")
```

### Checking for No Asthma frequencies
```{r}
#subset dataset with only asthma people
noasthmaonly <- subset(analytic, ASTHMA4 != 1)
table(noasthmaonly $ASTHMA4)
nrow(noasthmaonly)

AsthmaFreq <- table(noasthmaonly$ASTHMA4)
AsthmaFreq
write.csv(AsthmaFreq, file = "Asthma.csv")

#USING MACROS
library(gtools)
#use defmacro to define the macro
FreqTbl <-defmacro(OutputTable, InputVar, CSVTable, 
expr={
OutputTable <- table(InputVar);
write.csv(OutputTable, file = paste0(CSVTable, ".csv"))
})

FreqTbl (AlcGrpFreq, noasthmaonly$ALCGRP, "Alc")
FreqTbl (AgeGrpFreq, noasthmaonly$X_AGE_G, "Age")
FreqTbl (SexFreq, noasthmaonly$SEX, "Sex")
FreqTbl (HispFreq, noasthmaonly$X_HISPANC, "Hisp")
FreqTbl (RaceFreq, noasthmaonly$RACEGRP, "Race")
FreqTbl (MaritalFreq, noasthmaonly$MARGRP, "Mar")
FreqTbl (EdFreq, noasthmaonly$EDGROUP, "Ed")
FreqTbl (IncFreq, noasthmaonly$INCOME3, "Inc")
FreqTbl (BMIFreq, noasthmaonly$BMICAT, "BMI")
FreqTbl (SmokeFreq, noasthmaonly$SMOKGRP, "Smok")
FreqTbl (ExerFreq, noasthmaonly$EXERANY3, "Exer")
FreqTbl (HlthPlanFreq, noasthmaonly$HLTHPLN2, "HlthPln")
FreqTbl (GenHlthFreq, noasthmaonly$GENHLTH2, "GenHlth")
```

### Means and Standard Deviations
```{r}
mean(analytic$SLEPTIM2)
sd(analytic$SLEPTIM2)

#load package plyr
library(plyr)
#example
ddply(analytic,~ALCGRP,summarise,mean=mean(SLEPTIM2),sd=sd(SLEPTIM2))

#USING MACROS
library(gtools)
SumTbl <- defmacro(OutputTable, GroupVar, CSVTable,
expr={
OutputTable <- ddply(analytic,~GroupVar,summarise,mean=mean(SLEPTIM2),sd=sd(SLEPTIM2));
write.csv(OutputTable, file = paste0(CSVTable, ".csv"))
})

SumTbl (AlcGrpSum, analytic$ALCGRP, "Alc")
SumTbl (AgeGrpSum, analytic$X_AGE_G, "Age")
SumTbl (SexSum, analytic$SEX, "Sex")
SumTbl (HispSum, analytic$X_HISPANC, "Hisp")
SumTbl (RaceSum, analytic$RACEGRP, "Race")
SumTbl (MaritalSum, analytic$MARGRP, "Mar")
SumTbl (EdSum, analytic$EDGROUP, "Ed")
SumTbl (IncSum, analytic$INCOME3, "Inc")
SumTbl (BMISum, analytic$BMICAT, "BMI")
SumTbl (SmokeSum, analytic$SMOKGRP, "Smok")
SumTbl (ExerSum, analytic$EXERANY3, "Exer")
SumTbl (HlthPlanSum, analytic$HLTHPLN2, "HlthPln")
SumTbl (GenHlthSum, analytic$GENHLTH2, "GenHlth")
```

###weights example
```{r}
WeightVarList <- c("X_STATE", "X_LLCPWT", "ASTHMA3")

BRFSS_weights <- subset(BRFSS_a[WeightVarList])

colnames(BRFSS_weights)
nrow(BRFSS_weights)

#use questionr package

library(questionr)

WeightedAsthma <- wtd.table(BRFSS_weights$ASTHMA3, 
	y=BRFSS_weights$X_STATE, weights = BRFSS_weights$X_LLCPWT, normwt = FALSE, na.rm = TRUE,
  	na.show = FALSE)
write.csv(WeightedAsthma, file = "WeightedAsthma.csv")
```

### Table1 Chisq
```{r}
#load MASS library

library(MASS)

#make table

AlcTbl = table(analytic$ASTHMA4, analytic$ALCGRP) 

#run test
chisq.test(AlcTbl)  

#make macro

library(gtools)

ChiTest <- defmacro(VarName, TblName, expr={
TblName = table(analytic$ASTHMA4, analytic$VarName); 
chisq.test(TblName)})

ChiTest(ALCGRP, AlcTbl)
ChiTest(X_AGE_G, AgeTbl)
ChiTest(SEX, SexTbl)
ChiTest(X_HISPANC, HispTbl)
ChiTest(RACEGRP, RaceTbl)
ChiTest(MARGRP, MarTbl)
ChiTest(EDGROUP, EdTbl)
ChiTest(INCOME3, IncTbl)
ChiTest(BMICAT, BMITbl)
ChiTest(SMOKGRP, SmokTbl)
ChiTest(EXERANY3, ExerTbl)
ChiTest(HLTHPLN2, HlthPlnTbl)
ChiTest(GENHLTH2, GenHlthTbl)
```

### ANOVAS for Table 1
```{r}
#example ANOVA

AlcANOVA <- lm(formula = SLEPTIM2 ~ ALCGRP, data = analytic)
summary(AlcANOVA)

#make macro

library(gtools)

ANOVATest <- defmacro(VarName, TblName, expr={
TblName<- lm(formula = SLEPTIM2 ~ VarName, data = analytic);
summary(TblName)})

#call macro

ANOVATest (ALCGRP, AlcANOVA)
ANOVATest (X_AGE_G, AgeANOVA)
ANOVATest (X_HISPANC, HispANOVA)
ANOVATest (RACEGRP, RaceANOVA)
ANOVATest (MARGRP, MarANOVA)
ANOVATest (EDGROUP, EdANOVA)
ANOVATest (INCOME3, IncANOVA)
ANOVATest (BMICAT, BMIANOVA)
ANOVATest (SMOKGRP, SmokANOVA)
ANOVATest (EXERANY3, ExerANOVA)
ANOVATest (HLTHPLN2, HlthPlnANOVA)
ANOVATest (GENHLTH2, GenHlthANOVA)
```

### ttests for Table 1
```{r}
t.test(analytic$SLEPTIM2~analytic$ASTHMA4)
t.test(analytic$SLEPTIM2~analytic$SEX)
```
### The complete Descriptive analysis of BRFFS data is done. We know that Descriptive analysis can lead to Regression. In the next part we will proceed with doing Linear Regression for this analysis.

